### 2020-09-24 Adaptation.Rmd
new

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(lsmeans)
library(car)
library(agricolae)
library(multcomp)
```

```{r include=FALSE}
Data=read_csv("Alldata.csv")
```

```{r include=FALSE}
Data <- Data %>% 
  filter(state != "havelock")
```

# Gender

```{r}
Data1 <- Data %>% 
  filter(!is.na(gender)) %>% 
  group_by(state, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally() %>% 
  mutate(Perc=(n/nn)*100) 
```

```{r}
GrantJune = data.frame(FEM = 33,
                     MALE = 38)
chisq.test(GrantJune)

GrantJuly = data.frame(FEM = 27,
                     MALE = 36)
chisq.test(GrantJuly)

ClayJuly = data.frame(FEM = 37,
                     MALE = 34)
chisq.test(ClayJuly)

ClayJune = data.frame(FEM = 33,
                     MALE = 37)
chisq.test(ClayJune)

HavelockJune = data.frame(FEM = 14,
                     MALE = 19)
chisq.test(HavelockJune)

MacombJuly = data.frame(FEM = 13,
                     MALE = 22)
chisq.test(MacombJuly)

MacombJune = data.frame(FEM = 36,
                     MALE = 34)
chisq.test(MacombJune)

ArlingtonJuly = data.frame(FEM = 30,
                     MALE = 39)
chisq.test(ArlingtonJuly)

ArlingtonJune = data.frame(FEM = 33,
                     MALE = 38)
chisq.test(ArlingtonJune)
```

```{r}
Data1$state <- factor(Data1$state, levels=c("grant", "clay center", "havelock", "macomb", "arlington"), labels=c("Grant", "Clay Center", "Lincoln", "Macomb", "Arlington"))
ggplot(Data1, aes(x=gender, y=Perc, fill=gender)) + geom_bar(stat="identity", width=0.3) + facet_wrap(planting~state, ncol=4) + labs(x="gender", y="Percentage (%)") + theme(legend.position = "none") + theme_solarized() + ggsave("gender.png", units="in", width=12, height=6, dpi=600)
```

# Resultado Geral

```{r}
Data$state <- factor(Data$state, levels=c("grant", "clay center", "havelock", "macomb", "arlington"), labels=c("Grant, NE", "Clay Center, NE", "Lincoln, NE", "Macomb, IL", "Arlington, WI"))
```

```{r echo=FALSE, warning=FALSE, width=12, height=6}
ggplot(Data, aes(x=crop, y=weight, fill=crop)) + geom_boxplot() + 
  #geom_violin() + 
  stat_summary(fun.y=mean, fill="red", geom="point", 
                           shape=21, size=3, show_guide = FALSE) + 
  scale_x_discrete(breaks=c(0, 50, 100, 150, 200, 250, 300, 350, 400, 450, 500, 550, 600, 650, 700, 750)) +
  facet_wrap(planting~state, ncol=5) + scale_fill_brewer(palette = "Set3") +
  theme_wsj() + theme(legend.position="none", strip.text = element_text(size=14, face="bold")) + labs(x="", y="Biomass (g)") +
  ggsave("Adaptation.png", units="in", width=12, height=6, dpi=600)
```

```{r}
model1=lmer(weight ~ crop * state * planting + (1|plant), data=Data)
#summary(model1)
plot(ranef(model1))
shapiro.test(Data$weight)
Anova(model1, test.statistic = "F", type = "II")
lsmeans(model1, ~ crop | planting | state, contr="pairwise", adjust="none")

```

# Arlington

```{r echo=FALSE}
Arlington <- Data %>% 
  filter(state == "arlington")

model2=lmer(weight ~ crop * planting + (1|plant), data=Arlington)
#summary(model1)
plot(ranef(model2))
shapiro.test(Arlington$weight)
Anova(model2, test.statistic = "F", type = "II")
lsmeans(model2, "crop", contr="pairwise", adjust="none")
lsmeans(model2, "planting", contr="pairwise", adjust="none")
```

# Macomb

```{r echo=FALSE}
Macomb <- Data %>% 
  filter(state == "macomb")

model3=lmer(weight ~ crop * planting + (1|plant), data=Macomb)
#summary(model1)
plot(ranef(model3))
shapiro.test(Macomb$weight)
Anova(model3, test.statistic = "F", type = "II")
lsmeans(model3, ~ crop | planting, contr="pairwise", adjust="none")
```

# Grant

```{r echo=FALSE}
Grant <- Data %>% 
  filter(state == "grant")

model4=lmer(weight ~ crop * planting + (1|plant), data=Grant)
#summary(model1)
plot(ranef(model4))
shapiro.test(Grant$weight)
A=Anova(model4, test.statistic = "F", type = "II")
Grantlsm<-lsmeans(model4, ~ crop | planting, contr="pairwise", adjust="none")
cld(Grantlsm, by="crop", Letters=letters, adjust="none", reversed=TRUE, sort=TRUE)

library(multcomp)
ph <- glht(A, linfct = mcp(ind = "Tukey"))
cld(ph)
```

# Clay Center

```{r echo=FALSE}
Clay <- Data %>% 
  filter(state == "clay center")

model5=lmer(weight ~ crop * planting + (1|plant), data=Clay)
#summary(model1)
plot(ranef(model5))
shapiro.test(Clay$weight)
Anova(model5, test.statistic = "F", type = "II")
lsmeans(model5, ~ crop | planting, contr="pairwise", adjust="none")
```

# Havelock

```{r echo=FALSE}
Havelock <- Data %>% 
  filter(state == "havelock")

model6=lmer(weight ~ crop * planting + (1|plant), data=Clay)
#summary(model1)
plot(ranef(model6))
shapiro.test(Havelock$weight)
Anova(model6, test.statistic = "F", type = "II")
lsmeans(model6, ~ crop | planting, contr="pairwise", adjust="none")
```