### 2020-09-24 Analysis.Rmd

---
title: "Results"
author: "Maxwel Coura Oliveira"
date: "10/3/2019"
output: pdf_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(ggpubr)
library(bestNormalize)
library(ggthemes)
library(nortest)
library(pbkrtest)
```

```{r include=FALSE}
data <- read_csv("Data.csv") %>% 
  drop_na(weight)
data$year <- factor(data$year)
glimpse(data)
```

```{r warning=FALSE}
data %>% 
#  filter(location == "clay center") %>% 
ggplot() + aes(x=doyh, y=weight, color=planting, size=height) + geom_jitter(alpha=0.5) +  facet_wrap(location~crop, ncol = 3, scales = "free_y") + theme_bw() + geom_smooth(method = "lm") +
  ggsave("biomass.pdf", height=18, width=9)
#  geom_vline(xintercept = 182, linetype="dotted", 
#                color = "blue", size=1.5)
```

# Testing Normality
```{r}
ggdensity(data$weight)
pearson.test(data$weight)
pearson.test(data$weight)
```

```{r}
ggdensity(log(data$weight))

pearson.test(log(data$weight))
pearson.test(data$weight) 
```

```{r}
databn <- bestNormalize(data$weight)
data$biomass <- databn$x.t
ggdensity(data$biomass)
```

```{r}
pearson.test(data$biomass)
pearson.test(data$biomass)
```

```{r}
ggplot(data, aes(x=crop, y=weight, color=crop)) + geom_jitter(alpha=0.4) +
  geom_boxplot() +
   coord_flip() + theme_bw() +
  facet_grid(year~planting~location) 
```

```{r warning=FALSE}
Model <- lmer(log(weight) ~ crop*planting*location + (1|year:plant), data=data)
summary(Model)
anova(Model)
```

```{r}
Model1 <- lmer(log(weight) ~ crop*planting + (1|year:plant), data=data)
```

```{r}
PBmodcomp(Model1, Model, nsim=500, cl=2) 
```

```{r}
anova(Model)
```

```{r}
lsm <- emmeans(Model, ~ crop * planting | location ,  adjust="none", type="response")
lsm
```

```{r warning=FALSE}
cld <- CLD(lsm, adjust="none", Letters = letters, reversed = TRUE, type="response")
cld

#write_csv(cld, path="cld.csv")
```

```{r}
cld$location <- factor(cld$location, levels = c("grant", "clay center",  "havelock", "macomb",  "arlington"), labels=c("Grant, NE", "Clay Center, NE",  "Lincoln, NE", "Macomb, IL", "Arlington, WI"))

cld$crop <- factor(cld$crop, levels = c("corn", "soybean", "fallow"), labels=c("corn", "soybean", "fallow"))
cld$planting <- factor(cld$planting, levels =c("June", "July"), labels=c("June", "July"))
```

```{r}
ggplot(cld, aes(x=crop, y=response, color=crop, label=.group)) + 
  geom_point(size=3) + theme_bw() + labs(x="", y=expression(paste("Plamer amaranth (g plant"^"-1",")"))) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.1, size=1) +
  geom_text(nudge_y = c(100, 80, 80, 70, 40, 30, 200, 170, 100, 70, 60, 50, 260, 100, 100, 
                        100, 80, 50, 140, 100, 90, 50, 60, 50, 100, 80, 70, 70, 60, 60), size=6) +
  facet_grid(planting~location) + coord_flip() +
  scale_color_manual(values = c("blue", "darkgreen", "black", "orange", "red")) +
  theme(legend.position = "none", strip.text = element_text(size=18, face="bold"), axis.title = element_text(size=18),
        axis.text.x = element_text(size=17, angle = 45, hjust = 1), axis.text.y = element_text(size=17)) +
  ggsave("adaptation.png", height=8, width=20, dpi=600)
```

```{r}
data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) 
```

```{r}
arlington1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "arlington" & planting == "June")

  chisq.test(arlington1$n)

arlington2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "arlington" & planting == "July")

  chisq.test(arlington2$n) 
```

```{r}
macomb1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "macomb" & planting == "June")

chisq.test(macomb1$n)

macomb2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "macomb" & planting == "July")

chisq.test(macomb2$n) 
```

```{r}
havelock1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "havelock" & planting == "June")

chisq.test(havelock1$n)

havelock2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "havelock" & planting == "July")

chisq.test(havelock2$n) 
```

```{r}
clay1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "clay center" & planting == "June")

chisq.test(clay1$n)

clay2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "clay center" & planting == "July")

chisq.test(clay2$n) 
```

```{r}
grant1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "grant" & planting == "June")

chisq.test(grant1$n)

grant2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "grant" & planting == "July")

chisq.test(grant2$n)
```