### 2021-06-07 new_analyses.Rmd
as

---
title: "new_analses"
author: "Maxwel Coura Oliveira"
date: "6/7/2021"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
data <- read_csv("Data.csv")
```

```{r}
data %>% 
  filter(location != "grant") %>% 
  mutate_if(is_character, as_factor) -> data1
```

```{r}
data1 %>% 
  ggplot(aes(x = crop, y = weight, color = crop)) +
  geom_boxplot() +
  geom_point() +
  facet_grid(location ~ planting)
```