### 2021-06-08 new_analyses.Rmd
as

---
title: "new_analses"
author: "Maxwel Coura Oliveira"
date: "6/7/2021"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(drc)
library(broom)
```

```{r}
data <- read_csv("data2.csv")
```

```{r}
data %>% 
  filter(location != "grant") -> data1
```

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting)) %>% 
  mutate_if(is_character, as_factor) -> data2
```

```{r}
data2 %>% 
  mutate(harvest = as.character(harvest)) %>% 
  mutate(harvest = str_replace(harvest, "2018", "2019"),
         harvest = ymd(harvest)) %>% 
  filter(!is.na(harvest)) %>% 
  dplyr::select(location, month, crop, harvest) %>% 
  group_by(location, month, crop) %>% 
  count(harvest, name = "harvest_day") -> harvest
```

```{r}
data2 %>% 
  filter(!is.na(harvest)) %>% 
  dplyr::select(location, month, crop, harvest) %>% 
  group_by(location, month, crop) %>% 
  count(location, month, crop, name = "total") -> total
```

```{r}
total %>% 
  left_join(harvest) %>% 
  mutate(prop = harvest_day / total) %>% 
  mutate(lag = lag(prop)) %>% 
  mutate(prob = cumsum(prop)) -> cum_sum
```

```{r}
data2 %>% 
  mutate(harvest = str_replace(harvest, "2018", "2019"),
         harvest = ymd(harvest)) %>% 
  filter(!is.na(harvest)) %>% 
  left_join(cum_sum) %>% 
  rename(harvest_com = harvest) %>% 
  dplyr::select(-lag) -> data3
```

```{r}
data3 %>% 
  ggplot(aes(x = dap, y = prob, color = crop)) +
  geom_point() +
  facet_grid(location ~ month)
```

# Arlington

```{r}
arlington <- data3 %>% 
  filter(location == "arlington") %>% 
  filter(month == "July")
```

```{r}
model_arl <- drm(prob*100 ~ doyh, crop, fct = W2.3(), data = arlington)

plot(model_arl, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
     xlab = "Day of the year")
```

```{r}
#broom::tidy(model_arl)
summary(model_arl)
```

```{r}
ED(model_arl, c(10, 50, 90), interval = "delta", type = "absolute")
```