### 2021-07-12 new_analyses.Rmd
,jkj

---
title: "new_analses"
author: "Maxwel Coura Oliveira"
date: "6/7/2021"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(drc)
library(broom)
library(tidytext)
library(ggthemes)
library(ggtext)
```

```{r}
data <- read_csv("data2.csv")
```

```{r}
data %>% 
  filter(location != "grant") %>% 
  mutate(location = str_replace(location, "havelock", "lincon"))-> data1
```

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting)) %>% 
  mutate_if(is_character, as_factor) -> data2
```

```{r}
data2 %>% 
  mutate(harvest = as.character(harvest)) %>% 
  mutate(harvest = str_replace(harvest, "2018", "2019"),
         harvest = ymd(harvest)) %>% 
  filter(!is.na(harvest)) %>% 
  dplyr::select(location, month, crop, harvest) %>% 
  group_by(location, month, crop) %>% 
  count(harvest, name = "harvest_day") -> harvest
```

```{r}
data2 %>% 
  filter(!is.na(harvest)) %>% 
  dplyr::select(location, month, crop, harvest) %>% 
  group_by(location, month, crop) %>% 
  count(location, month, crop, name = "total") -> total
```

```{r}
total %>% 
  left_join(harvest) %>% 
  mutate(prop = harvest_day / total) %>% 
  mutate(lag = lag(prop)) %>% 
  mutate(prob = cumsum(prop)) -> cum_sum
```

```{r}
data2 %>% 
  mutate(harvest = str_replace(harvest, "2018", "2019"),
         harvest = ymd(harvest)) %>% 
  filter(!is.na(harvest)) %>% 
  left_join(cum_sum) %>% 
  rename(harvest_com = harvest) %>% 
  dplyr::select(-lag) -> data3
```

```{r}
data3 %>% 
  ggplot(aes(x = dap, y = prob, color = crop)) +
  geom_point() +
  facet_grid(location ~ month)
```
```{r}
car::leveneTest(prob ~ location, data = data3)
```

```{r}
data3 %>% 
  unite("trt", c("location", "crop"), sep = "-") -> data4
```

# June

```{r}
june <-  data4 %>% 
  filter(month == "June")
model <- drm(prob*100 ~ doyh, trt, fct = W2.3(fixed = c(NA, 100, NA)), data = june)

plot(model, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
     xlab = "Day of the year")
```
```{r}
#broom::tidy(model_arl)
summary(model)
```

```{r}
ed_values <- ED(model, c(10, 50, 90), interval = "delta", type = "absolute")
```

```{r}
ed_values %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "name") %>% 
  as_tibble() %>% 
  separate(name, c("param", "trt", "ED"), sep = ":") %>% 
  janitor::clean_names() %>% 
  separate("trt", c("location", "crop"), sep = "-") %>% 
  mutate(location = str_to_title(location)) %>% 
  mutate_if(is_character, as_factor) -> ed_values1
```

```{r}
label1 <- tibble(ed = "50% flowering", crop = "corn", 
                   x = 3.5, y = 185, label = "Transplanting \nday")
rect1 <- tibble(ed = "50% flowering", crop = "corn", 
               x = 3.5, y = 153, xend = 3.5, yend = 165)
```

```{r}
month <- tibble(ed = "90% flowering",
                crop = "corn",
                x = c(0.65, 0.65, 0.65, 0.65),
                y = c(168, 197, 228.5, 259), 
                label = c("June", "July", "August", "September"))

ed_values1 %>% 
  filter(crop == "corn") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(crop,  ed) %>% 
  mutate(location = reorder_within(location, 
                                   by = estimate, 
                                   within = ed)) %>% 
  ggplot(aes(x = location, y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 153, ymax= 183), 
            fill = "#FFFDD0", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 153, size = 1.1, 
             color = "red",
             alpha = 1) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2, color = c("#8B8000", "#80461b", 
                                  "darkgreen", "blue")) +
  labs(y = NULL, x = NULL,
       title = NULL) +
  facet_grid(ed ~ crop, scales = "free_y") +
  geom_curve(data = rect1, aes(x = x, y = y, 
                               xend = xend, yend = yend),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) + 
  geom_text(data = label1, 
            aes(x = x, y = y), 
            size = 2,
            label = label1$label, color = "red") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        plot.title = element_markdown(),
        strip.text = element_markdown(face = "bold"),
        plot.title.position = "plot") -> p1c
```

```{r}
month <- tibble(ed = "90% flowering",
                crop = "fallow",
                x = c(0.65, 0.65, 0.65, 0.65),
                y = c(168, 197, 228.5, 259), 
                label = c("June", "July", "August", "September"))

ed_values1 %>% 
  filter(crop == "fallow") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(crop,  ed) %>% 
  mutate(location = reorder_within(location, 
                                   by = estimate, 
                                   within = ed)) %>% 
  ggplot(aes(x = location, y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 153, ymax= 183), 
            fill = "#FFFDD0", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 153, size = 1.1, 
             color = "red",
             alpha = 1) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2, color = c("#8B8000", "#80461b", 
                                  "darkgreen", "blue")) +
  labs(y = "Day of year", x = NULL,
       title = NULL) +
  facet_grid(ed ~ crop, scales = "free_y") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        plot.title = element_markdown(),
        strip.text = element_markdown(face = "bold"),
        plot.title.position = "plot") -> p1f
```

```{r}
month <- tibble(ed = "90% flowering",
                crop = "corn",
                x = c(0.65, 0.65, 0.65, 0.65),
                y = c(168, 197, 228.5, 259), 
                label = c("June", "July", "August", "September"))

ed_values1 %>% 
  filter(crop == "corn") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(crop,  ed) %>% 
  mutate(location = reorder_within(location, 
                                   by = estimate, 
                                   within = ed)) %>% 
  ggplot(aes(x = location, y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 153, ymax= 183), 
            fill = "#FFFDD0", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 153, size = 1.1, 
             color = "red",
             alpha = 1) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2, color = c("#8B8000", "#80461b", 
                                  "darkgreen", "blue")) +
  labs(y = NULL, x = NULL,
       title = NULL) +
  facet_grid(ed ~ crop, scales = "free_y") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        plot.title = element_markdown(),
        strip.text = element_markdown(face = "bold"),
        plot.title.position = "plot") -> p1s
```

```{r}
library(patchwork)
f1 <- p1c | p1f | p1s +
  plot_annotation(
  title = "DOY of 10, 50, and 90% Palmer amaranth flowering after first transplanting") 
  ggsave("June.png", width = 9, height = 5, dpi = 600, f1)
```

# July

```{r}
july <-  data4 %>% 
  filter(month == "July")
model2 <- drm(prob*100 ~ doyh, trt, fct = W2.3(fixed = c(NA, 100, NA)), data = july)

plot(model2, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
     xlab = "Day of the year")
```

```{r}
#broom::tidy(model_arl)
summary(model2)
```

```{r}
ed_values2 <- ED(model2, c(10, 50, 90), interval = "delta", type = "absolute")
```

```{r}
ed_values2 %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "name") %>% 
  as_tibble() %>% 
  separate(name, c("param", "trt", "ED"), sep = ":") %>% 
  janitor::clean_names() %>% 
  separate("trt", c("location", "crop"), sep = "-") %>% 
  mutate(location = str_to_title(location)) %>% 
  mutate_if(is_character, as_factor) -> ed_values3
```

```{r}
label2 <- tibble(ed = "50% flowering", crop = "corn", 
                   x = 3.5, y = 211, label = "Transplanting \nday")
rect2 <- tibble(ed = "50% flowering", crop = "corn", 
               x = 3.5, y = 183, xend = 3.5, yend = 195)
```

```{r}
month <- tibble(ed = "90% flowering",
                crop = "corn",
                x = c( 0.65, 0.65, 0.65),
                y = c(197, 228.5, 259), 
                label = c("July", "August", "September"))
```

```{r}
ed_values3 %>% 
  filter(crop == "corn") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(crop,  ed) %>% 
  mutate(location = reorder_within(location, 
                                   by = estimate, 
                                   within = ed)) %>%
  ggplot(aes(x = location,
             y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 183, size = 1.1, 
             color = "red",
             alpha = 1) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2.5, color = c("#80461b", "darkgreen", "blue")) +
  labs(y = NULL, x = NULL) +
  facet_grid(ed ~ crop, scales = "free_y") +
  geom_curve(data = rect2, aes(x = x, y = y, 
                               xend = xend, yend = yend),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) + 
  geom_text(data = label2, 
            aes(x = x, y = y), 
            size = 2,
            label = label2$label, color = "red") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold"),
        plot.title = element_markdown(),
        plot.title.position = "plot") -> p2c
```

```{r}
month <- tibble(ed = "90% flowering",
                crop = "fallow",
                x = c( 0.65, 0.65, 0.65),
                y = c(197, 228.5, 259), 
                label = c("July", "August", "September"))
```

```{r}
ed_values3 %>% 
  filter(crop == "fallow") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(crop,  ed) %>% 
  mutate(location = reorder_within(location, 
                                   by = estimate, 
                                   within = ed)) %>%
  ggplot(aes(x = location,
             y = estimate)) +
    scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 183, size = 1.1, 
             color = "red",
             alpha = 1) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2.5, color = c("#80461b", "darkgreen", "blue")) +
  labs(y = "Day of year", x = NULL) +
  facet_grid(ed ~ crop, scales = "free_y") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold"),
        plot.title = element_markdown(),
        plot.title.position = "plot") -> p2f
```

```{r}
month <- tibble(ed = "90% flowering",
                crop = "soybean",
                x = c( 0.65, 0.65, 0.65),
                y = c(197, 228.5, 259), 
                label = c("July", "August", "September"))

ed_values3 %>% 
  filter(crop == "soybean") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(crop,  ed) %>% 
  mutate(location = reorder_within(location, 
                                   by = estimate, 
                                   within = ed)) %>%
  ggplot(aes(x = location,
             y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 183, size = 1.1, 
             color = "red",
             alpha = 1) +
  geom_point(size = 2) +
  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2.5, color = c("#80461b", "darkgreen", "blue")) +
  labs(y = NULL, x = NULL) +
  facet_grid(ed ~ crop, scales = "free_y") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold"),
        plot.title = element_markdown(),
        plot.title.position = "plot") -> p2s
p2s
```

```{r}
library(patchwork)
f3 <- p2c | p2f | p2s +
  plot_annotation(
  title = "DOY of 10, 50, and 90% Palmer amaranth flowering after second transplanting") 
  ggsave("July.png", width = 9, height = 5, dpi = 600, f3)
```

#### -----

```{r}
library(ggdist)
library(PupillometryR)
```

# Biomass

```{r}
data <- read_csv("data2.csv")
```

```{r}
data %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting)) %>% 
  mutate_if(is_character, as_factor) -> data1
```

```{r}
data1 %>% 
  mutate(month = fct_recode(month,
                            "first planting" = "June",
                            "second planting" = "July")) -> data1
```

```{r}
data1 %>% 
  mutate(location = str_to_title(location)) %>% 
  filter(crop == "corn") %>% 
  filter(!is.na(weight)) %>% 
  ggplot(aes(x = fct_reorder(location, weight), y = weight,
             color = location, 
             fill = location)) + 
  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   stat = "ydensity",
                   alpha = 0.8, scale = "count") +
  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  theme_test() +
  facet_grid(month ~ crop, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold")) -> fg1
```

```{r}
data1 %>% 
  mutate(location = str_to_title(location)) %>% 
  filter(crop == "fallow") %>% 
  filter(!is.na(weight)) %>% 
  ggplot(aes(x = fct_reorder(location, weight), y = weight,
             color = location, 
             fill = location)) + 
  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   stat = "ydensity",
                   alpha = 0.8, scale = "count") +
  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  theme_test() +
  facet_grid(month ~ crop, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = "Biomass (g)") +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold")) -> fg2
```

```{r}
data1 %>% 
  mutate(location = str_to_title(location)) %>% 
  filter(crop == "soybean") %>% 
  filter(!is.na(weight)) %>% 
  ggplot(aes(x = fct_reorder(location, weight), y = weight,
             color = location, 
             fill = location)) + 
  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   stat = "ydensity",
                   alpha = 0.8, scale = "count") +
  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  theme_test() +
  facet_grid(month ~ crop, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold")) -> fg3
```

```{r}
f4 <- fg1 | fg2 | fg3 +
  plot_annotation(
  title = NULL) 
  ggsave("biomass.png", width = 9, height = 5, dpi = 600, f4)
```

# Height

```{r}
data1 %>% 
  mutate(location = str_to_title(location)) %>% 
  filter(crop == "corn") %>% 
  filter(!is.na(height)) %>% 
  ggplot(aes(x = fct_reorder(location, height), y = height,
             color = location, 
             fill = location)) + 
  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   stat = "ydensity",
                   alpha = 0.8, scale = "count") +
  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  theme_test() +
  facet_grid(month ~ crop, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold")) -> fg4
```

```{r}
data1 %>% 
  mutate(location = str_to_title(location)) %>% 
  filter(crop == "fallow") %>% 
  filter(!is.na(height)) %>% 
  ggplot(aes(x = fct_reorder(location, height), y = height,
             color = location, 
             fill = location)) + 
  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   stat = "ydensity",
                   alpha = 0.8, scale = "count") +
  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  theme_test() +
  facet_grid(month ~ crop, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = "Height (cm)") +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold")) -> fg5
```

```{r}
data1 %>% 
  mutate(location = str_to_title(location)) %>% 
  filter(crop == "soybean") %>% 
  filter(!is.na(height)) %>% 
  ggplot(aes(x = fct_reorder(location, height), y = height,
             color = location, 
             fill = location)) + 
  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   stat = "ydensity",
                   alpha = 0.8, scale = "count") +
  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
  scale_color_viridis_d(option = "H") + 
  scale_fill_viridis_d(option = "H") + 
  theme_test() +
  facet_grid(month ~ crop, scales = "free") +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold")) -> fg6
```

```{r}
f5 <- fg4 | fg5 | fg6 +
  plot_annotation(
  title = NULL) 
  ggsave("height.png", width = 9, height = 5, dpi = 300, f5)
```

###--------------------------

```{r}
data %>% 
  filter(location != "grant") %>% 
  mutate(location = str_replace(location, "havelock", "lincon"))-> data1
```

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting)) %>% 
  mutate_if(is_character, as_factor) -> data2
```

```{r}
data2 %>% 
  mutate(harvest = as.character(harvest)) %>% 
  mutate(harvest = str_replace(harvest, "2018", "2019"),
         harvest = ymd(harvest)) %>% 
  filter(!is.na(harvest)) %>% 
  dplyr::select(location, month, crop, harvest) %>% 
  group_by(month, crop) %>% 
  count(harvest, name = "harvest_day") -> harvest
```

```{r}
data2 %>% 
  filter(!is.na(harvest)) %>% 
  dplyr::select(location, month, crop, harvest) %>% 
  group_by(month, crop) %>% 
  count(month, crop, name = "total") -> total
```

```{r}
total %>% 
  left_join(harvest) %>% 
  mutate(prop = harvest_day / total) %>% 
  mutate(lag = lag(prop)) %>% 
  mutate(prob = cumsum(prop)) %>% 
  mutate(doyh = yday(harvest)) -> cum_sum
```

```{r}
data2 %>% 
  mutate(harvest = str_replace(harvest, "2018", "2019"),
         harvest = ymd(harvest)) %>% 
  filter(!is.na(harvest)) %>% 
  ungroup() %>% 
  left_join(cum_sum) %>% 
  rename(harvest_com = harvest) %>% 
  dplyr::select(-lag) -> data3
```

# June

```{r}
data3 %>% 
  filter(month == "June") -> month_jun 
model3 <- drm(prob*100 ~ doyh, crop, fct = W1.3(), data = month_jun)

plot(model3, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
     xlab = "Day of the year")
```

```{r}
#broom::tidy(model_arl)
summary(model3)
```

```{r}
ed_values3 <- ED(model3, c(10, 50, 90), interval = "delta", type = "absolute")
```

```{r}
ed_values3 %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "name") %>% 
  as_tibble() %>% 
  separate(name, c("param", "crop", "ED"), sep = ":") %>% 
  janitor::clean_names() %>% 
  mutate_if(is_character, as_factor) -> ed_values3
```

```{r}
label1 <- tibble(ed = "90% flowering", 
                   x = 2.5, y = 185, label = "Transplanting \nday")
rect1 <- tibble(ed = "90% flowering", 
               x = 2.5, y = 153, xend = 2.5, yend = 165)
```

```{r}
month <- tibble(ed = "90% flowering",
                x = c(0.65, 0.65, 0.65, 0.65),
                y = c(168, 197, 228.5, 259), 
                label = c("June", "July", "August", "September"))
```

```{r}
ed_values3 %>% 
  mutate(month = "First transplanting") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(month,  ed) %>% 
  mutate(crop = reorder_within(crop, 
                                   by = estimate, 
                                   within = ed)) %>% 
  ggplot(aes(x = crop, y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 153, ymax= 183), 
            fill = "#FFFDD0", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 153, size = 1.1, 
             color = "#CCCC00",
             alpha = 1) +
  geom_point(size = 2, aes(color = crop, 
                           shape = crop)) +
  geom_linerange(aes(ymin = lower, ymax = upper, color = crop, 
                     shape = crop), size = 1.3) +
  scale_color_manual(values = c("#412234","#EE4B2B","#32CD32",
                                "#412234","#EE4B2B","#32CD32",
                                "#412234","#EE4B2B","#32CD32")) +
  scale_shape_manual(values = c(15,16,17,15,16,17,15,16,17)) +
  facet_grid(ed ~ month, scales = "free_y") +
  coord_flip() +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2, color = c("#8B8000", "#80461b", 
                                  "darkgreen", "blue")) +
  labs(y = NULL, x = NULL,
       title = NULL) +
  geom_curve(data = rect1, aes(x = x, y = y, 
                               xend = xend, yend = yend),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) + 
  geom_text(data = label1, 
            aes(x = x, y = y), 
            size = 2,
            label = label1$label, color = "#CCCC00") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        plot.title = element_markdown(),
        strip.text = element_markdown(face = "bold"),
        plot.title.position = "plot") -> p1c
p1c
```

# July

```{r}
data3 %>% 
  filter(month == "July") -> month_jul 
model4 <- drm(prob*100 ~ doyh, crop, fct = W1.3(), data = month_jul)

plot(model4, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
     xlab = "Day of the year")
```

```{r}
#broom::tidy(model_arl)
summary(model4)
```

```{r}
ed_values4 <- ED(model4, c(10, 50, 90), interval = "delta", type = "absolute")
```

```{r}
ed_values4 %>% 
  as.data.frame() %>% 
  rownames_to_column(., var = "name") %>% 
  as_tibble() %>% 
  separate(name, c("param", "crop", "ED"), sep = ":") %>% 
  janitor::clean_names() %>% 
  mutate_if(is_character, as_factor) -> ed_values4
```

```{r}
label2 <- tibble(ed = "90% flowering", 
                   x = 2.5, y = 211, label = "Transplanting \nday")
rect2 <- tibble(ed = "90% flowering",  
               x = 2.5, y = 183, xend = 2.5, yend = 195)
```

```{r}
month1 <- tibble(ed = "90% flowering",
                x = c(0.65, 0.65, 0.65),
                y = c(197, 228.5, 259), 
                label = c("July", "August", "September"))
```

```{r}
ed_values4 %>% 
  mutate(month = "Second transplanting") %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  group_by(month,  ed) %>% 
  mutate(crop = reorder_within(crop, 
                                   by = estimate, 
                                   within = ed)) %>% 
  ggplot(aes(x = crop, y = estimate)) +
  scale_x_reordered(expand = c(.01, .01)) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 183, ymax= 214), 
            fill = "#80461b", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 214, ymax= 244), 
            fill = "lightgreen", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
                        ymin= 244, ymax= 274), 
            fill = "blue", alpha=0.05) +
  geom_hline(yintercept = 183, size = 1.1, 
             color = "#D2691E",
             alpha = 1) +
  geom_point(size = 2, aes(color = crop, shape = crop)) +
  geom_linerange(aes(ymin = lower, ymax = upper, color = crop), size = 1.3) +
  scale_color_manual(values = c("#412234","#EE4B2B","#32CD32",
                                "#412234","#EE4B2B","#32CD32",
                                "#412234","#32CD32","#EE4B2B")) +
  scale_shape_manual(values = c(15,16,17,15,16,17,15,17,16)) +
  geom_text(data = month1, aes(x = x, y = y), label = month1$label,
            size = 2, color = c("#80461b", "darkgreen", "blue")) +
  labs(y = NULL, x = NULL) +
  facet_grid(ed ~ month, scales = "free_y") +
  geom_curve(data = rect2, aes(x = x, y = y, 
                               xend = xend, yend = yend),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) + 
  geom_text(data = label2, 
            aes(x = x, y = y), 
            size = 2,
            label = label2$label, color = "#D2691E") +
  coord_flip() +
  theme_test() +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold"),
        plot.title = element_markdown(),
        plot.title.position = "plot") -> p2c
```

```{r}
library(patchwork)
f3 <- p1c | p2c +
  plot_annotation(
  title = "DOY of 10, 50, and 90% Palmer amaranth flowering") 
  ggsave("Combined.png", width = 6, height = 5, dpi = 600, f3)
```

# Biomass

```{r}
data <- read_csv("data2.csv")
```

```{r}
data %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting)) -> data1
```

```{r}
data1 %>% 
  mutate(month = fct_recode(month,
                            "First transplanting" = "June",
                            "Second transplanting" = "July")) -> data1
```

```{r}
data1 %>% 
  filter(!is.na(weight)) %>% 
  group_by(location, month, crop, doyh) %>% 
  summarise(biomass = sum(weight))  %>% 
  ggplot(aes(x = biomass, y = crop)) +
  geom_density_ridges() +
  facet_grid(location ~ month)
```

```{r}
options(scipen = 999)
library(ggstream)
library(ggridges)

data1 %>% 
  filter(!is.na(weight)) %>% 
  dplyr::select(month, crop, weight, doyh) %>% 
  group_by(month, crop) %>% 
  mutate(n = n()) %>% 
  mutate(rate = weight / n) %>% 
  group_by(month, crop, doyh) %>% 
  summarise(biomass = sum(rate)) %>% 
  mutate(doyh = as.integer(doyh)) -> biomass
```

```{r}
biomass %>% 
  ggplot(aes(x = doyh, y = biomass, fill = crop, label = crop)) +
  geom_stream(bw = 0.75, geom = "polygon", type = "ridge") +
#  geom_stream_label(size = 4, type = "proportional", n_grid = 1000) +
  facet_grid(~ month, scales = "free_x") +
  scale_fill_manual(values = c("#412234","#EE4B2B","#32CD32")) +
  labs(y = "Biomass (g plant)") +
  theme_test()

ggsave("biomass.png")
```

```{r}
options(scipen = 999)
library(ggstream)
library(ggridges)

data1 %>% 
  filter(!is.na(height)) %>% 
  dplyr::select(month, crop, height, doyh) %>% 
  group_by(month, crop) %>% 
  mutate(n = n()) %>% 
  mutate(rate = height / n) %>% 
  group_by(month, crop, doyh) %>% 
  summarise(height = sum(rate)) %>% 
  mutate(doyh = as.integer(doyh)) -> height
```

```{r}
height %>% 
  ggplot(aes(x = doyh, y = height, fill = crop, label = crop)) +
  geom_stream(bw = 0.75, geom = "polygon", type = "proportional") +
  facet_grid(~ month, scales = "free_x") +
  geom_stream_label(size = 4, type = "proportional", n_grid = 1000) +
  scale_fill_manual(values = c("#412234","#EE4B2B","#32CD32")) +
  labs(y = "Height (g plant)") +
  theme_test()
```