### 2021-07-20 data analysis/map.Rmd

---
title: "map"
author: "Maxwel Coura Oliveira"
date: "2/1/2019"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)
library(ggthemes)
library(usmap)
library(ggmap)
library(ggrepel)
library(maps)
library(grid)
library(mapdata)
library(maptools)
library(stringr)
library(dplyr)
library(ggsn)
library(choroplethr)
```

```{r include=FALSE, warning=FALSE}
region <-c("Arlington, WI", "Macomb, IL", "Grant, NE", "Lincoln, NE", "Clay Center, NE")
Latitude <-c(43.335148, 40.45921,  40.83999664, 40.806862, 40.51999792)
Longitude <-c(-89.372523, -90.6718, -101.722330444, -96.681679, -98.052999788)

Map<-data.frame(region, Latitude, Longitude)
```

```{r include=FALSE}
states = map_data("state")
df <- subset(states, region %in% c("wisconsin", "nebraska", "illinois", "iowa"))
```

```{r include=FALSE}
counties <- map_data("county")
midwest <- subset(counties, region %in% c("wisconsin", "nebraska", "illinois", "iowa"))
```

```{r map, echo=FALSE}
ggplot(data = df, mapping = aes(x = long, y = lat, group = group, color=region)) + 
  geom_point(data=Map, mapping=aes(x=Longitude, y=Latitude, group=region, color=region), size=5) +
  geom_polygon(data = midwest, fill = "NA", color = "gray") +
  geom_polygon(color = "black", fill = NA) + scale_color_brewer(palette=c("Set1")) +
  labs(fill="", y="", x="", caption="") +
  north(df, location = 'topright', symbol = 1, anchor = c(x = -86, y = 47.3)) +
  theme_map() + 
  scalebar(data=df, dist = 100, dist_unit="km", model = 'International', location = "bottomleft", transform = TRUE) + 
  annotate("text", x = -101, y = 38, label = "Figure: Maxwel C Oliveira", size=4) +
      theme(axis.text=element_blank(), 
        axis.title=element_blank(), 
        panel.background = element_rect(fill="white", color = "black"), 
        panel.grid.major = element_line(color = "white"), 
  panel.grid.minor = element_blank(),
  legend.position=c(0.05,0.8), legend.text = element_text(size = 20, colour = "black"), legend.key = element_rect(fill=NA), panel.border = element_rect(fill=NA,color="black", size=0.5, linetype="solid"),
legend.background = element_rect(fill =NA), legend.key.width = unit(2.2, "line"), legend.title=element_blank()) +
  ggsave("map.png", dpi=600, width=18, height=12)
```