### 2021-08-11 data analysis/weight.Rmd

---
title: "weight"
author: "Maxwel Coura Oliveira"
date: "8/11/2021"
output: html_document
---

###------------
```{r}
library(tidyverse)
library(lubridate)
library(lme4)
library(lmerTest)
library(emmeans)
library(tidytext)
library(ggthemes)
library(ggtext)
library(patchwork)
library(extrafont)
library(showtext)
showtext_auto()
font_add_google("Roboto", "roboto") 
```

```{r}
data <- read_csv("../data/data2.csv")
```

###--------------------------

```{r}
data %>% 
  mutate(location = str_replace(location, "havelock", "lincoln")) %>% 
  mutate(harvest = case_when(
    location == "clay center" & doyh == 169 ~ "8/18/2018",
    TRUE ~ harvest
  )) -> data1
```

```{r}
data1 %>% 
  rename(month = planting) %>% 
  mutate_at(c("year", "location"), as_factor) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) -> data2
```

```{r}
model <- lmer(log(weight) ~ crop * month + 
                                 (1|location/year), data = data2)
```

```{r}
anova(model)
```

```{r}
emmeans(model, ~ crop * month, type = "response") -> emmeans1
```

```{r}
multcomp::cld(emmeans1, 
              alpha=0.05, 
              Letters=letters, 
              adjust="none", reversed = TRUE) %>% 
  as_tibble() -> weight
```

```{r}
weight %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>%  
  ggplot(aes(y = crop, 
             x = response,
             shape = month,
             color = month)) +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  geom_jitter(data = data2 %>% 
                rename(response = weight) %>% 
                mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")),
#                filter(month == "First transplanting"), 
              alpha = 0.01, aes(color = month, shape = month), size = 1,
              position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_color_manual(values = c(2,1)) +
  geom_text(aes(label = round(response,1), x = upper.CL + 130, color = month),
            size = 8, fontface = "bold",
            position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_point(size = 2,
             position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_x_continuous(limits = c(0,1201), breaks = seq(0, 1200, 300)) +
  scale_shape_manual(values = c(1,16)) +
  labs(x = expression(paste("Weight (g plant "^"-1",")")), 
       y = NULL, color = NULL) +
  annotate("text", x = 300, y = 1.7, 
           label = "First cohort",
           size = 7, color = 1,
           fontface = "bold") + 
  annotate("text", x = 300, y = 2.3, 
           label = "Second cohort",
           size = 7, color = 2,
           fontface = "bold") + 
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> figA

#ggsave("figures/biomass.png", width = 4, height = 4)
```

```{r}
model2 = lmer(log(height) ~ crop * month + 
                                 (1|location/year), data = data2)
anova(model2)
```

```{r}
emmeans(model2, ~ crop * month, type = "response", cont="pairwise", adjust="none") -> emmeans2
```

```{r}
multcomp::cld(emmeans2$emmeans, 
              alpha=0.05, 
              Letters=letters, 
              adjust="none", reversed = TRUE) %>% 
  as_tibble() -> height
```

```{r warning = FALSE}
height %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>%
  ggplot(aes(y = crop,
             x = response,
             shape = month,
             color = month)) +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  geom_jitter(data = data2 %>% 
                rename(response = height) %>% 
                mutate(crop = fct_recode(crop,
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")),
#                filter(month == "First transplanting"), 
              alpha = 0.02, aes(color = month, shape = month), size = 1,
              position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_color_manual(values = c(2,1)) +
  geom_text(aes(label = round(response,1), x = upper.CL + 20, color = month),
            size = 8, fontface = "bold",
            position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_point(size = 2,
             position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_shape_manual(values = c(1,16)) +
#  scale_x_continuous(limits = c(0,300), breaks = c(0,25,50,75,100)) +
  labs(x = expression(paste("Height (cm plant "^"-1",")")), 
       y = NULL, color = NULL) +
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> figB

#ggsave("figures/height.png", width = 4, height = 4)
```

```{r}
(figA | figB) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))

ggsave("figures/Figure 3.png", width = 6, height = 3)
```

```{r eval=FALSE}
wrap_elements(full = curve1) / (figure) / (figA | figB) +
   plot_layout(heights = c(1.5, 1.25, 1.25))

#ggsave("figures/combine 3.png", height = 9)
```

```{r}
library(emoGG)
```

```{r}
data2 %>% 
  mutate(gender_A = case_when(
    gender == "m" ~ 0,
    gender == "f" ~ 1,
    TRUE ~ NA_real_
  )) %>% 
  mutate_if(is_character, as.factor) %>% 
  filter(!is.na(gender_A)) %>% 
  filter(!is.na(doyh)) %>% 
  filter(!is.na(weight)) %>% 
  filter(!is.na(height)) %>% 
  mutate(doyh = as.numeric(doyh)) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) -> data3
```

```{r}
first <- tibble(crop = "bareground",
                x = 200, y = 0.73,
                label = "First cohort")
second <- tibble(crop = "bareground",
                 x = 235, y = 0.4,
                label = "Second cohort")
cohortf1 <- tibble(crop = "corn",
                 x = 215, y = 0.94)
cohortm1 <- tibble(crop = "corn",
                 x = 215, y = 0.06)
```

```{r}
data3 %>% 
  ggplot(aes(x = doyh, y = gender_A)) +
  geom_point(alpha = 0.1, aes(color = month)) +
  geom_emoji(data = cohortf1, aes(x = x, y = y), emoji = "2640") +
  geom_emoji(data = cohortm1, aes(x = x, y = y), emoji = "2642") +
  geom_smooth(aes(color = month), 
              method = "gam", 
              se = TRUE, 
              method.args=list(family="binomial", 
                               type = "response", 
                               trans = "plogis"),
              stat = "smooth") +
  theme_test(base_family = "roboto") +
  facet_grid(~crop) +
  scale_color_manual(values = c(2,1)) +
  scale_fill_manual(values = c(2,1)) +
  scale_x_continuous(limits = c(170, 260), 
                     breaks = seq(170, 260, 15)) +
  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
  labs(x = "Day of year", y = NULL) +
  geom_text(data = first, mapping = aes(x = x, y = y), label = first$label,
            size = 10, color = 1) +
  geom_text(data = second, mapping = aes(x = x, y = y), label = second$label,
            size = 10, color = 2) +
  theme(legend.position = "none",
        axis.title = element_markdown(size = 35),
        axis.text = element_markdown(size = 25),
        strip.text = element_markdown(face = "bold", size = 30)) -> pl_doyh
#  xlim(175, 270)

ggsave("figures/fig.png", width = 9, height = 4)
```

```{r}
cohortf2 <- tibble(crop = "corn",
                 x = 92, y = 0.94)
cohortm2 <- tibble(crop = "corn",
                 x = 92, y = 0.06)
```

```{r}
data3 %>% 
  mutate(crop = fct_relevel(crop,
                            levels = c("bareground", "corn", "soybean"))) %>% 
  ggplot(aes(x = weight, y = gender_A)) +
  geom_point(alpha = 0.1, aes(color = month)) +
  geom_emoji(data = cohortf2, aes(x = x, y = y), emoji = "2640") +
  geom_emoji(data = cohortm2, aes(x = x, y = y), emoji = "2642") +
  geom_smooth(aes(color = month), 
              method = "gam", 
              se = FALSE, 
              stat = "smooth") +
  theme_test(base_family = "roboto") +
  facet_grid(~crop, scales = "free") +
  scale_color_manual(values = c(2,1)) +
  scale_fill_manual(values = c(2,1)) +
  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
  labs(x = expression(paste("Weight (g plant "^"-1",")")), y = NULL) +
  theme(legend.position = "none",
        axis.title = element_markdown(size = 35),
        axis.text = element_markdown(size = 25),
        strip.text = element_markdown(face = "bold", size = 30)) -> pl_weight
#  xlim(175, 270)

ggsave("fig.png", width = 9, height = 4)
```
```{r}
cohortf2 <- tibble(crop = "corn",
                 x = 110, y = 0.94)
cohortm2 <- tibble(crop = "corn",
                 x = 110, y = 0.06)
```

```{r}
data3 %>% 
  mutate(crop = fct_relevel(crop,
                            levels = c("bareground", "corn", "soybean"))) %>% 
  ggplot(aes(x = height, y = gender_A)) +
  geom_point(alpha = 0.1, aes(color = month)) +
  geom_emoji(data = cohortf2, aes(x = x, y = y), emoji = "2640") +
  geom_emoji(data = cohortm2, aes(x = x, y = y), emoji = "2642") +
  geom_smooth(aes(color = month), 
              method = "gam", 
              se = FALSE, 
              stat = "smooth") +
  theme_test(base_family = "roboto") +
  facet_grid(~crop) +
  scale_color_manual(values = c(2,1)) +
  scale_fill_manual(values = c(2,1)) +
  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
  labs(x = expression(paste("Height (cm plant "^"-1",")")), y = NULL) +
    theme(legend.position = "none",
        axis.title = element_markdown(size = 35),
        axis.text = element_markdown(size = 25),
        strip.text = element_markdown(face = "bold", size = 30)) -> pl_height
#  xlim(175, 270)

ggsave("fig.png", width = 9, height = 4)
```

```{r}
pl_doyh / pl_weight / pl_height

ggsave("figures/Figure 4.png", width = 6, height = 8)
```