### 2021-08-13 data analysis/weather/weather.Rmd

---
title: "weather"
author: "Maxwel Coura Oliveira"
date: "7/8/2021"
output: html_document
---

```{r}
library(tidyverse) # Data wrangling
library(lubridate) # Dates operations
library(daymetr)
library(chirps)
library(patchwork)
library(ggtext)
library(nasapower)
library(vegan)
```

# Load data

You can load your data with read_csv if you have it in a spreadsheet.

```{r}
# your data
df_time <- tibble(ID = c('1','2','3','4','5'), 
                  Site = c('Arlington','Macomb','Lincoln',
                               'Clay Center', 'Grant'),
                  # Both coordinates in decimal format
                  latitude = c(43.18, 40.45, 40.85, 40.57, 40.8430),
                  longitude = c(-89.20, -90.67, -96.62, -98.13, -101.72),
                  Start = c('1988-01-01','1988-01-01','1988-01-01',
                                '1988-01-01', '1988-01-01'),
                  Early_planting = c("2018-06-01", "2018-06-01",
                                     "2018-06-01",
                                     "2018-06-01", "2018-06-01"),
                  Late_planting = c("2018-06-01", "2018-06-01", "2018-06-01",
                                    "2018-06-01", "2018-06-01"),
                      End = c('2020-12-31','2020-12-31','2020-12-31',
                              '2020-12-31', '2020-12-31'))

df_time <- df_time %>% 
  mutate_at(5:8, ymd) # as date
```

## Daymet

```{r function}
source("weather_daymet.R")
```

```{r get_data}
#dpp default = 0 - dpp = days prior start date

df_daymet <- weather_daymet(input = df_time, dpp = 0) 
```

```{r}
df_daymet %>% 
  janitor::clean_names() %>% 
  mutate(month_lab = month(date, label = TRUE)) -> adaptation_weather
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) -> adaptation_weather1
```

```{r}
adaptation_weather1 %>% 
  filter(year %in% c(2018, 2019)) %>% 
  group_by(site, year, month) %>% 
  summarise(prec = sum(pp), tmax = mean(tmax), tmin = mean(tmin), tmean = mean(tmean)) %>% 
  mutate_if(is.double, ~round(., 1))
```

```{r}
adaptation_weather %>% 
  mutate(site = fct_recode(site, 
                           "Arlington, WI" = "Arlington",
                           "Clay Center, NE" = "Clay Center",
                           "Grant, NE" = "Grant",
                           "Lincoln, NE" = "Lincoln",
                           "Macomb, IL" = "Macomb")) -> adaptation_weather
```

```{r}
text1 <- tibble(site = "Arlington, WI")
text2 <- tibble(site = "Clay Center, NE")
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) %>% 
  ggplot(aes(x = month_lab, y = tmean)) +
  geom_jitter(aes(color = variable), alpha = 0.05, shape = 0) +
  scale_color_manual(values = c("red", "blue", "gray90")) +
  geom_line(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean, group = 1), 
             color = "red") +
  geom_line(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean, group = 1), 
             color = "blue") +
  geom_point(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean), 
             fill = "red", shape = 22, size = 3) +
  geom_point(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean), 
             fill = "blue", shape = 22, size = 3) +
  geom_richtext(data = text1, aes(x = 2, y = 7), lineheight = 0.4, 
            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
            label.color = NA, fill = NA) +
  geom_richtext(data = text2, aes(x = 2.2, y = 10), lineheight = 0.4, 
            label = "30 years <br>daily mean <br>air temperature", size = 6, 
             label.color = NA, fill = NA, color = "gray90") +
  geom_curve(data = text2, 
             x = 2, y = 13, 
             xend = 2.5, yend = 16,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = -0.3, color = "gray90") + 
  scale_y_continuous(limits = c(0, 32)) +
  facet_grid(~ site) +
  labs(x = NULL, y = "Temperature (C)") +
  theme_test() +
  theme(legend.position = "none",
        strip.text = element_markdown(size = 28),
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 27),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25))  -> temp

ggsave("weather.png", height = 3, width = 6)
```

```{r}
text3 <- tibble(site = "Grant, NE")
text4 <- tibble(site = "Lincoln, NE")
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  group_by(month_lab, site, year) %>% 
  summarise(pp = sum(pp)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) -> adaptation_pp
```

```{r message = FALSE, warning = FALSE}
adaptation_pp %>% 
ggplot(aes(x = month_lab, y = pp)) +
  geom_bar(data = adaptation_pp %>% filter(year == c(2018, 2019)), 
           mapping = aes(fill = year),
           stat = "identity", position = "dodge", alpha = 1) +
  geom_point(data = adaptation_pp %>% filter(year != c(2018, 2019)), 
              position = position_jitter(seed = 1),
              alpha = 0.2, color = "gray50", 
              show.legend = FALSE) +
  scale_fill_manual(values = c("red", "blue")) +
  facet_grid(~ site) +
  geom_richtext(data = text3, aes(x = 2.6, y = 330), lineheight = 0.4, 
            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
            label.color = NA, fill = NA) +
  geom_richtext(data = text4, aes(x = 3.6, y = 320), lineheight = 0.4, 
            label = "30 years <br>monthly sum <br>precipitation", size = 6, 
             label.color = NA, fill = NA, color = "gray90") +
  geom_curve(data = text4, 
             x = 2.6, y = 310, 
             xend = 2.1, yend = 320,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = -0.3, color = "gray90") + 
  labs(x = NULL, y = "Precipitation (mm)") +
  theme_test() +
  theme(legend.position = "none",
        strip.text = element_markdown(size = 28),
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 27),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> ppp

```

```{r}
temp / ppp

ggsave("Figure 1.png")
```