### 2021-08-16 data analysis/layout.Rmd
b

---
title: "layout"
author: "Maxwel Coura Oliveira"
date: "8/16/2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggimage)
library(ggtext)
```

```{r}
read_csv("../data/layout.csv") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  mutate(distance = as.double(distance)) %>% 
  mutate(emoji = case_when(
    crop == "corn" ~ "https://img.icons8.com/color/48/000000/corn.png",
    crop == "soybean" ~ "https://img.icons8.com/color/48/000000/soy.png",
    TRUE ~ NA_character_)) %>% 
  mutate(weed = "https://hotemoji.com/images/emoji/0/1gqjxj81p5mvj0.png") -> layout
```

```{r}
layout %>% 
  ggplot(aes(x = row, y = distance)) +
  facet_grid(~crop) +
#  geom_point() +
#  scale_x_discrete(limits = c(-0.5, 4.5)) +
  geom_image(data = layout,
               aes(image=weed)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 4.5, image=emoji)) +
  labs(x= NULL, y = NULL) +
  theme_test() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        strip.text = element_markdown(size = 18),
        panel.background = element_rect(fill = "#f7e4d7")) -> fig1
```

```{r}
library(ggimage)
fig1 +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 4.5, image=emoji))

ggsave("figures/layout.png", height = 4, width = 9)
```