### 2021-08-17 data analysis/layout.Rmd
gh

---
title: "layout"
author: "Maxwel Coura Oliveira"
date: "8/16/2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggimage)
library(ggtext)
```

```{r}
read_csv("../data/layout.csv") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  mutate(distance = as.double(distance)) %>% 
  mutate(emoji = case_when(
    crop == "corn" ~ "https://img.icons8.com/color/48/000000/corn.png",
    crop == "soybean" ~ "https://img.icons8.com/color/48/000000/soy.png",
    TRUE ~ NA_character_)) %>% 
  mutate(weed = "https://hotemoji.com/images/emoji/0/1gqjxj81p5mvj0.png") -> layout
```

```{r}
layout %>% 
  ggplot(aes(x = row, y = distance)) +
  facet_grid(~crop) +
#  geom_point() +
#  scale_x_discrete(limits = c(-0.5, 4.5)) +
  geom_image(data = layout,
               aes(image=weed)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 4.5, image=emoji)) +
  labs(x= NULL, y = NULL) +
  theme_test() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        strip.text = element_markdown(size = 18),
        panel.background = element_rect(fill = "#f7e4d7")) -> fig1
```

```{r}
library(ggimage)
fig1 +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 4.5, image=emoji)) -> fig2
```

```{r}
palmer <- tibble(crop = "bareground")
corn <- tibble(crop = "corn")
soybean <- tibble(crop = "soybean")
```

```{r}
fig2 +
  geom_text(data = palmer, aes(x = 2.7, y = 270), size = 2, hjust = 0, 
            fontface = "italic",
            label = "24 Palmer amaranth plants \nwere placed 76.2 cm \nat the same distance") +
  geom_text(data = palmer, aes(x = 2.3, y = 228), size = 2,
            label = "76.2 cm", fontface = "bold") +
  geom_text(data = palmer, aes(x = 2.5, y = 390), size = 2,
            label = "Palmer amaranth \nplants", fontface = "bold") +
  geom_text(data = corn, aes(x = 2.5, y = 400), size = 2,
            label = "Corn plants", fontface = "bold") +
  geom_text(data = soybean, aes(x = 2.5, y = 400), size = 2,
            label = "Soybean plants", fontface = "bold") -> fig3

#ggsave("figures/layout.png", height = 4, width = 9)
```

```{r}
fig3 +
  geom_segment(data = palmer, mapping = aes(x = 2.1, y = 264, 
                                            xend = 2.85, yend = 230),
               linetype = "dotted") +
  geom_segment(data = palmer, mapping = aes(x = 2.1, y = 188, 
                                            xend = 2.85, yend = 228),
               linetype = "dotted") +
  geom_segment(data = palmer, mapping = aes(x = 2, y = 250, 
                                            xend = 2, yend = 198),
               linetype = "dotted") -> fig4

#ggsave("figures/layout.png", height = 4, width = 9)
```

```{r}
fig4 +
  geom_curve(data = palmer, aes(x = 2.5, y = 405, 
                               xend = 2.1, yend = 420),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) +
  geom_curve(data = palmer, aes(x = 2.5, y = 375, 
                               xend = 2.85, yend = 375),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) +
  geom_curve(data = corn, aes(x = 2.3, y = 405, 
                               xend = 2.4, yend = 420),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3) +
  geom_curve(data = corn, aes(x = 2.7, y = 395, 
                               xend = 2.6, yend = 380),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3) + 
  geom_curve(data = soybean, aes(x = 2.3, y = 405, 
                               xend = 2.4, yend = 420),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3) +
  geom_curve(data = soybean, aes(x = 2.7, y = 395, 
                               xend = 2.6, yend = 380),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3)

ggsave("figures/layout.png", height = 4, width = 9)
```